---
title: "Owens Illinois"
author: "Pallav Routh"
output: html_document
---

**Load Libraries**

```{r,warning=FALSE,message=FALSE}
library(readxl)
library(ggplot2)
library(dplyr)
```

Experimenting with Uganda only.

**Read in the data set**

```{r}
uganda <- read_excel("C:/Users/routh/Desktop/Study Materials/Projects/O-I/Uganda/uganda.xlsx",col_types = c("date", "numeric", "text", "text", "numeric", "numeric"))
```

**Changing variable names**

```{r}
glimpse(uganda)
#rename the col headers
colnames(uganda)[3] <- "product"
colnames(uganda)[5] <- "tonnes"
colnames(uganda)[1] <- "date"
colnames(uganda)[4] <- "customer"
```

**clean up the customer name colomn**

```{r}
customer <- uganda$customer 
customerID <- sapply(strsplit(customer," "), `[`, 1)
customerID[1:30]
uganda <- cbind(uganda,customerID)
```

What are the overall product types with count?  
What are the time stamps and the counts?

```{r}
table(as.factor(uganda$product))
table(as.factor(uganda$date))
```

The table above suggests that we might be dealing with monthly totals.The time stamp suggests that we cannot break down the months to weeks. We have to use monthly totals as time series data points. This might pose a problem since it gives us only 52 data points.

How has the average product shipment tonnes change over the years for Uganda?

```{r}
uganda <- uganda%>%
                mutate(Year = format(date, "%Y"), Month = format(date, "%m"))

#temp1 <- uganda%>%
               #group_by(product,Year,Month)%>%
               #summarise(mean_tonnes = mean(tonnes))%>%
               #filter(mean_tonnes != 0)
#temp1   what this was doing is taking only the first months of every year. We need to first find the monthly total for each cactegory every year

#g1 <- ggplot(temp1,aes(Year,mean_tonnes,fill=factor(product)))
#g1 + geom_bar(stat="identity",position="dodge")+scale_fill_discrete(name = "product")

temp2 <- uganda%>%
               arrange(Year, product, Month)%>%    #only if you want to view this in the specific way
               group_by(product, Year,  Month)%>%
               summarise(monthly_sum=sum(tonnes))%>%
               group_by(product)%>%
               mutate(month_index = row_number())

g2 <- ggplot(temp2, aes(month_index, monthly_sum, colour = product))
g2+geom_line()

temp3 <- temp2%>%
              group_by(product,Year)%>%
              summarise(mean_monthly_tonnes = mean(monthly_sum))

g3 <- ggplot(temp3,aes(Year,mean_monthly_tonnes,fill=factor(product)))
g3 + geom_bar(stat="identity",position="dodge")+scale_fill_discrete(name = "product")
```

One way to check the demand of a particular product is to look at how many unique customers there are every year

```{r}
temp4 <- uganda%>%
               group_by(product,Year)%>%
               summarise(unique_customers = n_distinct(customerID))

temp4

temp5 <- left_join(temp3,temp4,by=c("product","Year"))

temp5
```

NAB, Spirits, Spirits and wine has done well in the average sales figures. 

Let's track a loyal customer and see if the trend persists.

```{r}
x <- data.frame(table(uganda$customer)) # customer 193 seems to be loyal 
temp6 <- uganda%>%
               filter(customerID == 193)%>%  #a spirit dealer
               group_by(Year,Month)%>%
               summarise(monthly_sum = sum(tonnes))%>%
               group_by(Year)%>%
               summarise(monthly_mean = mean(monthly_sum))

temp6

temp7 <- uganda%>%
               filter(customerID == 219)%>%  #another spirit dealer
               group_by(Year,Month)%>%
               summarise(monthly_sum = sum(tonnes))%>%
               group_by(Year)%>%
               summarise(monthly_mean = mean(monthly_sum))
temp7

# looks like spirit dealers are the 'loyal customers'
```

Looks like spirit dealers are the 'loyal customers'. Let's see if this hunch is true. Let's see if the customers who had the highest number of transactions are the ones who deal in spirit. Look at temp5. Only spirits have their sales going up along with the number of unique customers. 

How many transactions are there every month for every product category every year?

```{r}
temp8 <- uganda%>%
               group_by(product,Year,Month)%>%
               summarise(n_transactions = n())

temp8
```

This tells us that there are possibly more than one transaction every day in some cases and no transaction every day. 


